---
title: "SURE Project"
author: ""
date: "`r format(Sys.Date(), '%B %d, %Y')`"
graphics: yes
output: pdf_document
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      cache = T,
                      eval = T,
                      message = F,
                      warning = F,
                      fig.width = 8,
                      fig.height = 10,
                      fig.align = 'center')
library(sp)
library(sf)
library(terra)
library(raster)
library(rgdal)
library(pander)
library(magrittr)
library(ggplot2)
library(dplyr)
library(tidyr)
library(RColorBrewer)
library(here)
library(plotrix)
```

## Objective

Compare spatial predictions of kelp to "in situ" survey data. Compare each year and location for 150, 300, 600, 900 resolutions.

## Extraction

Extract the predicted $log$ kelps density of every year (2004 - 2021) for each site in the North Coast.

```{r}
# set a directory
w.dir <- here()
d.dir <- here('data')
r.dir <- '/Volumes/Chunting HD/Git_Repositories/Chunting_Spatial_Analyses/spatial_data'

r1.dir <- paste(r.dir, 'sp_predictions_300m', sep ='/')
r2.dir <- paste(r.dir, 'sp_predictions_1500m', sep ='/')
r3.dir <- paste(r.dir, 'sp_predictions_600m', sep ='/')
r4.dir <- paste(r.dir, 'sp_predictions_900m', sep ='/')
r5.dir <- paste(r.dir, 'sp_predictions_150m', sep ='/')
r6.dir <- paste(r.dir, 'sp_predictions_120m', sep ='/')

# read and transform the observed data to the log scale
df <- read.csv(paste(d.dir, 
                     'RCCA_kelp_inverts_NC_depth-zones_wave_clim_temp_nit_subs_orbvel_npp.csv', 
                     sep = '/')) %>%
  dplyr::select(site_name, year, transect, zone, latitude, longitude, den_NERLUE) %>%
  mutate_at(vars(year, transect, zone, site_name), list(as.factor)) %>%
  mutate(log_den_NERLUE = log(den_NERLUE)) 
head(df) 
```

Note that $\log(0)$ returns -Inf. How to deal with $\log(0)$?

```{r}
df$log_den_NERLUE <- replace(df$log_den_NERLUE, df$log_den_NERLUE == -Inf, 0)
```

Calculate the mean and standard error of kelps density of every year for each site by zone (INNER/OUTER).

```{r}
obs <- df %>% 
  group_by(site_name, year, zone) %>%
  summarise_at(vars(log_den_NERLUE), list(mean = mean, se = std.error), na.rm = TRUE) %>%
  pivot_wider(names_from = zone, values_from = c(mean, se))
head(obs)
```

Extract the predicted $log$ kelps density of every year for each site at different resolutions.

```{r}
# kelp density predictions at different resolutions ----

# # read the .csv file
# site <- read.csv(paste(d.dir, 'RCCA_North_Coast_sites.csv', sep = '/'))
# # convert from .csv to .shp
# site_shp <- st_as_sf(site, coords = c('longitude', 'latitude'), crs = 'EPSG:4326')
# 
# # declaring an empty data frame
# pred <- data.frame(site_name = character(),
#                    year = numeric(),
#                    fit = numeric())
# 
# for (i in c(2006:2021)) {
#   rast <- rast(paste0(r6.dir, paste0('/', i, '_Log_Nereo_NC.tif')))
#   ext <- terra::extract(rast, vect(site_shp$geometry)) %>%
#     mutate(site_name = site$site_name, year = as.factor(i), .before = fit) %>%
#     dplyr::select(-ID)
#   pred <- rbind(pred, ext)
# }
# 
# head(pred)
# 
# # write to cvs
# merge_df <- left_join(pred,
#                       site %>% dplyr::select(c(site_name, longitude, latitude)),
#                       by = 'site_name')
# 
# write.csv(merge_df, file.path(d.dir, 'NC_kelp_density_predictions_120m.csv'), row.names = FALSE)
```

## Comparison

```{r}
# kelp density predictions at 300m resolution
pred_300m <- read.csv(paste(d.dir, 'NC_kelp_density_predictions_300m.csv', sep ='/')) %>%
  mutate_at(vars(year, site_name), list(as.factor))

kelp_data_300m <- left_join(pred_300m, obs, by = c('site_name', 'year')) %>%
  group_by(site_name) %>%
  arrange(year, .by_group = TRUE) %>%
  relocate(fit, .after = last_col())
head(kelp_data_300m)
```

```{r}
# kelp density predictions at 1500m resolution
pred_1500m <- read.csv(paste(d.dir, 'NC_kelp_density_predictions_1500m.csv', sep ='/')) %>%
  mutate_at(vars(year, site_name), list(as.factor))

kelp_data_1500m <- left_join(pred_1500m, obs, by = c('site_name', 'year')) %>%
  group_by(site_name) %>%
  arrange(year, .by_group = TRUE) %>%
  relocate(fit, .after = last_col())
head(kelp_data_1500m)
```

## Plotting

Plot log of kelps density vs year for each site at different resolutions.

```{r}
sites <- unique(kelp_data_300m$site_name) 

kelp_longer_300m <- kelp_data_300m %>%
  dplyr::select(-c(longitude, latitude)) %>%
  pivot_longer(
    -c('site_name', 'year', 'fit'),
    names_to = c('.value', 'zone'),
    names_sep = '_'
    ) %>%
  mutate(resolution = as.factor(300))

kelp_longer_1500m <- kelp_data_1500m %>%
  dplyr::select(-c(longitude, latitude)) %>%
  pivot_longer(
    -c('site_name', 'year', 'fit'),
    names_to = c('.value', 'zone'),
    names_sep = '_'
    ) %>%
  mutate(resolution = as.factor(1500))

kelp_longer <- rbind(kelp_longer_300m, kelp_longer_1500m)

for (i in sites) {
  plot <- kelp_longer %>% 
    filter(site_name == i) %>%
    ggplot() + 
    geom_pointrange(aes(
      x = year, y = mean, group = zone, color = zone, 
      ymin = mean - se, ymax = mean + se
      ), alpha = 0.5, size = 0.3) + 
    geom_bar(aes(x = year, y = fit, 
                 fill = ifelse(!is.na(fit) & fit >= 6.6, 'YES', 'NO')), 
             stat = 'identity', position = 'dodge', alpha = 0.2) +
    facet_wrap(. ~ resolution, nrow = 3) + 
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, size = 8), 
          plot.title = element_text(hjust = 0.5),
          panel.grid.major = element_blank(),
          legend.title = element_text(size = 9),
          legend.text = element_text(size = 7)) + 
    labs(y = 'log of kelp density', title = i, fill = 'fit >= 6.6')
  print(plot)
}
```
