---
title: "SURE Project"
author: ""
date: "`r format(Sys.Date(), '%B %d, %Y')`"
graphics: yes
output: pdf_document
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      cache = T,
                      eval = T,
                      message = F,
                      warning = F,
                      fig.width = 5,
                      fig.height = 3,
                      fig.align = 'center')
library(sp)
library(sf)
library(terra)
library(raster)
library(rgdal)
library(pander)
library(magrittr)
library(ggplot2)
library(dplyr)
library(tidyr)
library(RColorBrewer)
library(here)
library(plotrix)
```

## Objective 

Compare spatial predictions of kelp to "in situ" survey data. Compare each year and location.

## Extraction

Extract the predicted $log$ kelps density of every year (2004 - 2021) for each site in the North Coast.

```{r}
# set a directory
w.dir <- here()
d.dir <- here('data')
r.dir <- here('spatial_data/sp_predictions_5.1.1_V2')
# rock.dir <- here('spatial_data/sp_predictions_5.1.1_V2_rock')

# read and transform the observed data to the log scale
df <- read.csv(paste(d.dir, 
                     'RCCA_kelp_inverts_NC_depth-zones_wave_clim_temp_nit_subs_orbvel_npp.csv', 
                     sep = '/')) %>%
  dplyr::select(site_name, year, transect, zone, latitude, longitude, den_NERLUE) %>%
  mutate_at(vars(year, transect, zone, site_name), list(as.factor)) %>%
  mutate(log_den_NERLUE = log(den_NERLUE)) 
head(df) 
```

Note that $\log(0)$ returns -Inf.

```{r}
filter(df, den_NERLUE == 0) %>% count() # 719 0's
```

How to deal with $\log(0)$?

```{r}
df$log_den_NERLUE <- replace(df$log_den_NERLUE, df$log_den_NERLUE == -Inf, 0)
head(df)
```

Calculate the mean and standard error of kelps density of every year for each site by zone (INNER/OUTER).

```{r}
obs <- df %>% 
  group_by(site_name, year, zone) %>%
  summarise_at(vars(log_den_NERLUE), list(mean = mean, se = std.error), na.rm = TRUE) %>%
  pivot_wider(names_from = zone, values_from = c(mean, se))
head(obs)
```

Extract the predicted $log$ kelps density of every year for each site.

```{r}
# read the .csv file
site <- read.csv(paste(d.dir, 
                       'RCCA_North_Coast_sites.csv', 
                       sep = '/')) 
# convert from .csv to .shp
site_shp <- st_as_sf(site, coords = c('longitude', 'latitude'), crs = 'EPSG:4326')
# write the file
st_write(site_shp, paste0(d.dir, '/RCCA_North_Coast_sites.shp'), append = FALSE)

# declaring an empty data frame 
pred <- data.frame(site_name = character(),
                   year = numeric(),
                   fit = numeric())

for (i in c(2006:2021)) {
  rast <- rast(paste0(r.dir, paste0('/', i, '_Nereo_preds_NC_V4_5.1.1_V2.tif')))
  ext <- terra::extract(rast, vect(site_shp$geometry)) %>%
    mutate(site_name = site$site_name, year = as.factor(i), .before = fit) %>%
    dplyr::select(-ID) 
  pred <- rbind(pred, ext)
}

head(pred)

# sites and years have predicted values higher than 6.6 (log of density)
filter(pred, fit >= 6.6)

# write to cvs
merge_df <- left_join(pred,
                      site %>% select(c(site_name, longitude, latitude)),
                      by = 'site_name')
write.csv(merge_df, file.path(d.dir, 'NC_kelp_density_predictions.csv'), row.names = FALSE)
```

## Comparison

Compare the predicted data to observed data.

```{r}
dim(obs)
dim(pred)

kelp_data <- left_join(pred, obs, by = c('site_name', 'year')) %>%
  group_by(site_name) %>%
  arrange(year, .by_group = TRUE) %>%
  relocate(fit, .after = last_col())
head(kelp_data)
```

## Plotting

Plot log of kelps density vs year for each site.

```{r}
# kelp_data %>% 
#   pivot_longer(
#     -c('site_name', 'year', 'fit'), 
#     names_to = c('.value', 'zone'),
#     names_sep = '_'
#     ) %>%
#   filter(site_name == 'Caspar') %>%
#   ggplot(aes(x = year, y = mean, fill = zone)) + 
#   geom_bar(position = 'dodge', stat = 'identity')

sites <- unique(kelp_data$site_name) 
kelp_longer <- kelp_data %>% 
    pivot_longer(
      -c('site_name', 'year', 'fit'),
      names_to = c('.value', 'zone'),
      names_sep = '_'
      ) 

kelp_longer %>% 
    filter(site_name == 'Caspar') %>%
    ggplot() + 
    geom_pointrange(aes(
      x = year, y = mean, group = zone, color = zone, 
      ymin = mean - se, ymax = mean + se
      ), alpha = 0.5, size = 0.3) + 
    geom_bar(aes(x = year, y = fit, fill = ifelse(fit >= 6.6, 'YES', 'NO')), 
             stat = 'identity', position = 'dodge', alpha = 0.2) + 
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, size = 8), 
          plot.title = element_text(hjust = 0.5),
          panel.grid.major = element_blank()) + 
    labs(y = 'log of kelp density', title = 'Glass Beach')
  
for (i in sites) {
  plot <- kelp_longer %>% 
    filter(site_name == i) %>%
    ggplot() + 
    geom_pointrange(aes(
      x = year, y = mean, group = zone, color = zone, 
      ymin = mean - se, ymax = mean + se
      ), alpha = 0.5, size = 0.3) + 
    geom_bar(aes(x = year, y = fit, 
                 fill = ifelse(!is.na(fit) & fit >= 6.6, 'YES', 'NO')), 
             stat = 'identity', position = 'dodge', alpha = 0.2) + 
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, size = 8), 
          plot.title = element_text(hjust = 0.5),
          panel.grid.major = element_blank(),
          legend.title = element_text(size = 9),
          legend.text = element_text(size = 7)) + 
    labs(y = 'log of kelp density', title = i, fill = 'fit >= 6.6')
  print(plot)
}
```

Plot obs. vs pred.

```{r, fig.width = 8, fig.height = 6}
kelp_longer %>%
  ggplot(aes(x = fit, y = mean, color = zone)) +
  geom_point(alpha = 0.6) + 
  geom_smooth(method = 'lm', alpha = 0.15, aes(fill = zone)) + 
  xlim(NA, 110) + 
  labs(x = 'pred', y = 'obs') + 
  facet_wrap(~zone, ncol = 2) +
  theme_bw()
```